Report on the Fourth Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE4)

### 2017-03-10 WSSSPE4_report/WSSSPE4_report.tex
added all but two lightning talks

@@

    \item \textbf{Jack Dongarra, Sven Hammarling, Nicholas J.\ Higham, Samuel
    D.\ Relton\textsuperscript{\textasteriskcentered}, Pedro Valero-Lara,
    and Mawussi Zounon: \textit{Creating a Standardised Set of Batched
    BLAS Routines}}~\cite{Dongarra:2016ws}.
    \note{Description}

    \item \textbf{Lukas Breitwieser\textsuperscript{\textasteriskcentered},
    Roman Bauer, Alberto Di Meglio, Leonard Johard,
    Marcus Kaiser, Marco Manca, Manuel Mazzara, Fons Rademakers,
    Max Talanov: \textit{The BioDynaMo Project}}~\cite{Bauer:2016ws}.
    \note{Description}

    \item \textbf{Aseel Aldabjan, Robert Haines, and
    Caroline Jay\textsuperscript{\textasteriskcentered}: \textit{How should we
    measure the relationship between code quality and software
    sustainability?}}~\cite{Aldabjan:2016ws}.
    \note{Description}

>>>>

    \item \textbf{Neil Chue Hong\textsuperscript{\textasteriskcentered}:
    \textit{Making it easier to understand research software impact}}~\cite{ChueHong:2016wsb}.
    Chue Hong proposed that the WSSSPE community collaboratively develop a framework
    to better understand (and measure the impact of) research software. This could
    be based on a workflow involving mining the literature for mentions of software,
    identifying linkages between software, and visualizing dependencies between
    software or between software and other research artifacts. Many tools already
    exist in this ecosystem, so Chue Hong proposed that a WSSSPE working group
    collect these tools, summarize use cases, identify missing pieces, and find
    low-hanging fruit to integrate tools or build new ones.

    \item \textbf{Willem Robert van Hage\textsuperscript{\textasteriskcentered},
    Jason Maassen and Rob van Nieuwpoort: \textit{Software Impact Measurement at the
    Netherlands eScience Center}}~\cite{vanHage:2016ws}.
    \note{Description}

    \item \textbf{Gabrielle Allen, Emily Chen\textsuperscript{\textasteriskcentered},
    Ray Idaszak, and Daniel S.\ Katz: \textit{Report on Software Metrics for
    Research Software}}~\cite{GAllen:2016wsb}.
    \note{Description}

@@

    \item \textbf{Jack Dongarra, Sven Hammarling, Nicholas J.\ Higham, Samuel
    D.\ Relton\textsuperscript{\textasteriskcentered}, Pedro Valero-Lara,
    and Mawussi Zounon: \textit{Creating a Standardised Set of Batched
    BLAS Routines}}~\cite{Dongarra:2016ws}.
    This talk described Batched BLAS (BBLAS), a new standard for computing multiple
    BLAS (Basic Linear Algebra Subproblems) in parallel. Unlike existing parallel
    BLAS algorithms provided by, e.g., Intel MKL, cuBLAS, or MAGMA, BBLAS is
    optimized for smaller matrices, which apply to a variety of problems in areas
    such as machine learning, multifrontal solvers, image processing, fluid
    dynamics, and astrophysics. BBLAS offers nearly a factor of two improvement
    in computing speed compared to non-batched operations. To develop a sustainable
    BBLAS library, the developers have focused on proposing a standard API,
    holding workshops between academics and vendors to reach consensus, and
    performing research into API design decisions on performance.

    \item \textbf{Lukas Breitwieser\textsuperscript{\textasteriskcentered},
    Roman Bauer, Alberto Di Meglio, Leonard Johard,
    Marcus Kaiser, Marco Manca, Manuel Mazzara, Fons Rademakers,
    Max Talanov: \textit{The BioDynaMo Project}}~\cite{Bauer:2016ws}.
    Breitwieser described the BioDynaMo project, a large-scale platform for
    biological simulations that provides access to computational resources while
    hiding the complexity of distributed computing. It also promotes computational
    reproducibility through shared open-access data. The project began as a
    modernization initiative porting an existing simulation software, Cx3D, from
    Java to C++, but also incorporated many modern software engineering practices
    such as hosting and version control on GitHub, automated tests and continuous
    integration via Travis-CI, code coverage reports, and an automated build system.
    In addition, over 1000 students participated in development via the Intel
    Modern Code Developer Challenge competition---the winner achieved a
    performance speedup of 320 times. The project plans to continue improving
    the development process with (e.g.) static code checkers and Git commit hooks,
    continued architecture redesign, and verification.

    \item \textbf{Aseel Aldabjan, Robert Haines, and
    Caroline Jay\textsuperscript{\textasteriskcentered}: \textit{How should we
    measure the relationship between code quality and software
    sustainability?}}~\cite{Aldabjan:2016ws}.
    This talk explored the relationship between code quality and the sustainability
    of research software; however, the latter has no concrete definition, thus making
    it challenging to provide guidance on how to achieve sustainability. The authors
    proposed a ``software sustainment'' metric based on the time difference between
    the most recent and initial commits\slash contributions to the codebase.
    Using this metric, they analyzed 3113 Java projects hosted on GitHub, created
    in 2009, and found that 22\% and 35\% had sustainment metrics of 0 and less
    than seven, respectively. Interestingly, the authors also found high
    code quality (measured using a variety of static analytic metrics) across
    all these projects, with little correlation to sustainability.
